---
title: "PCM_S12"
author: "Marcial Escudero"
date: "2024-01-22"
output: html_document
editor_options: 
  chunk_output_type: console
---

Análisis Biogeográfico con BioGeoBEARS

A continuación, realizaremos los análisis biogeográficos utilizando el clado de Psychotria de Hawái (Ree & Smith, 2008). Evaluaremos y compararemos los siguientes modelos de evolución de rangos geográficos:

    DEC y DEC+J

    DIVALIKE y DIVALIKE+J

    BAYAREALIKE y BAYAREALIKE+J

Para cada uno de estos modelos, generaremos las reconstrucciones de los estados ancestrales y evaluaremos su ajuste estadístico a los datos.


```{r}

################################################################################
# BIOGEOBEARS ANALYSIS: M0_UNCONSTRAINED
# Models: DEC, DEC+J, DIVALIKE, DIVALIKE+J, BAYAREALIKE, BAYAREALIKE+J
################################################################################

# ------------------------------------------------------------------------------
# 1. SETUP AND DATA LOADING
# ------------------------------------------------------------------------------

# Load required libraries
library(rexpokit)
library(cladoRcpp)
library(ape)
library(BioGeoBEARS)

# Set working directory (Update this path to your actual working directory)
wd <- normalizePath("./")
setwd(wd)

# Define file paths for the phylogeny and geography data
# Example files are based on Hawaiian Psychotria (Ree & Smith 2008)
trfn   <- normalizePath("./Psychotria_5.2.newick")
geogfn <- normalizePath("./Psychotria_geog.data")

# Read and inspect the phylogeny and geographic range data
tr <- read.tree(trfn)
tipranges <- getranges_from_LagrangePHYLIP(lgdata_fn = geogfn)

# Define maximum range size (maximum number of areas a species can occupy)
max_range_size <- 4

# Location of BioGeoBEARS plot scripts
scriptdir <- normalizePath(system.file("extdata/a_scripts", package="BioGeoBEARS"))


# ------------------------------------------------------------------------------
# HELPER FUNCTION FOR CLEAN INITIALIZATION
# ------------------------------------------------------------------------------
# This function generates a fresh, clean BioGeoBEARS run object to avoid S4 class 
# inheritance errors when copying objects between models.
setup_base_bgb <- function(tree_file, geog_file, max_size) {
  bgb <- define_BioGeoBEARS_run()
  bgb$trfn <- tree_file
  bgb$geogfn <- geog_file
  bgb$max_range_size <- max_size
  bgb$speedup <- TRUE
  bgb$use_optimx <- TRUE
  bgb$num_cores_to_use <- 1
  bgb$force_sparse <- FALSE
  bgb$calc_ancprobs <- TRUE
  bgb$return_condlikes_table <- TRUE
  bgb$calc_TTL_loglike_from_condlikes_table <- TRUE
  bgb <- readfiles_BioGeoBEARS_run(bgb)
  return(bgb)
}


# ------------------------------------------------------------------------------
# 2. DEC & DEC+J MODELS
# ------------------------------------------------------------------------------

# --- Run DEC ---
bgb_dec <- setup_base_bgb(trfn, geogfn, max_range_size)
check_BioGeoBEARS_run(bgb_dec)

resfn_dec <- "Psychotria_DEC_M0_unconstrained_v1.Rdata"
if (!file.exists(resfn_dec)) {
  resDEC <- bears_optim_run(bgb_dec)
  save(resDEC, file = resfn_dec)
} else {
  load(resfn_dec)
}

# --- Run DEC+J ---
bgb_decj <- setup_base_bgb(trfn, geogfn, max_range_size) # Fresh initialization

# Use DEC estimates as starting values for DEC+J
dstart <- resDEC$outputs@params_table["d","est"]
estart <- resDEC$outputs@params_table["e","est"]

bgb_decj$BioGeoBEARS_model_object@params_table["d", c("init", "est")] <- dstart
bgb_decj$BioGeoBEARS_model_object@params_table["e", c("init", "est")] <- estart

# Add jump dispersal (j) as a free parameter
bgb_decj$BioGeoBEARS_model_object@params_table["j", "type"] <- "free"
bgb_decj$BioGeoBEARS_model_object@params_table["j", c("init", "est")] <- 0.0001

check_BioGeoBEARS_run(bgb_decj)

resfn_decj <- "Psychotria_DEC+J_M0_unconstrained_v1.Rdata"
if (!file.exists(resfn_decj)) {
  resDECj <- bears_optim_run(bgb_decj)
  save(resDECj, file = resfn_decj)
} else {
  load(resfn_decj)
}

# --- Plot DEC & DEC+J ---
pdf("Psychotria_DEC_vs_DEC+J_M0_unconstrained_v1.pdf", width=8.5, height=11)
plot_BioGeoBEARS_results(resDEC, "BioGeoBEARS DEC on Psychotria M0_unconstrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDEC, "BioGeoBEARS DEC on Psychotria M0_unconstrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDECj, "BioGeoBEARS DEC+J on Psychotria M0_unconstrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDECj, "BioGeoBEARS DEC+J on Psychotria M0_unconstrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
dev.off()

# ------------------------------------------------------------------------------
# 3. DIVALIKE & DIVALIKE+J MODELS
# ------------------------------------------------------------------------------

# --- Run DIVALIKE ---
bgb_diva <- setup_base_bgb(trfn, geogfn, max_range_size) 

# Set up DIVALIKE specific parameters (Remove subset-sympatry)
bgb_diva$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_diva$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_diva$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_diva$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "2-j"
bgb_diva$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/2"
bgb_diva$BioGeoBEARS_model_object@params_table["y", "type"] <- "ysv*1/2"
bgb_diva$BioGeoBEARS_model_object@params_table["v", "type"] <- "ysv*1/2"

# Allow classic widespread vicariance
bgb_diva$BioGeoBEARS_model_object@params_table["mx01v", "type"] <- "fixed"
bgb_diva$BioGeoBEARS_model_object@params_table["mx01v", "init"] <- 0.5
bgb_diva$BioGeoBEARS_model_object@params_table["mx01v", "est"] <- 0.5

check_BioGeoBEARS_run(bgb_diva)

resfn_diva <- "Psychotria_DIVALIKE_M0_unconstrained_v1.Rdata"
if (!file.exists(resfn_diva)) {
  resDIVALIKE <- bears_optim_run(bgb_diva)
  save(resDIVALIKE, file = resfn_diva)
} else {
  load(resfn_diva)
}

# --- Run DIVALIKE+J ---
bgb_divaj <- setup_base_bgb(trfn, geogfn, max_range_size) 

# Use DIVALIKE estimates as starting values
dstart <- resDIVALIKE$outputs@params_table["d","est"]
estart <- resDIVALIKE$outputs@params_table["e","est"]
bgb_divaj$BioGeoBEARS_model_object@params_table["d", "init"] <- dstart
bgb_divaj$BioGeoBEARS_model_object@params_table["d", "est"] <- dstart
bgb_divaj$BioGeoBEARS_model_object@params_table["e", "init"] <- estart
bgb_divaj$BioGeoBEARS_model_object@params_table["e", "est"] <- estart

# Set up DIVALIKE+J specific parameters
bgb_divaj$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_divaj$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_divaj$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_divaj$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "2-j"
bgb_divaj$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/2"
bgb_divaj$BioGeoBEARS_model_object@params_table["y", "type"] <- "ysv*1/2"
bgb_divaj$BioGeoBEARS_model_object@params_table["v", "type"] <- "ysv*1/2"

bgb_divaj$BioGeoBEARS_model_object@params_table["mx01v", "type"] <- "fixed"
bgb_divaj$BioGeoBEARS_model_object@params_table["mx01v", "init"] <- 0.5
bgb_divaj$BioGeoBEARS_model_object@params_table["mx01v", "est"] <- 0.5

# Add jump dispersal (j)
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "type"] <- "free"
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "init"] <- 0.0001
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "est"] <- 0.0001
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "min"] <- 0.00001
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "max"] <- 1.99999

check_BioGeoBEARS_run(bgb_divaj)

resfn_divaj <- "Psychotria_DIVALIKE+J_M0_unconstrained_v1.Rdata"
if (!file.exists(resfn_divaj)) {
  resDIVALIKEj <- bears_optim_run(bgb_divaj)
  save(resDIVALIKEj, file = resfn_divaj)
} else {
  load(resfn_divaj)
}

# --- Plot DIVALIKE & DIVALIKE+J ---
pdf("Psychotria_DIVALIKE_vs_DIVALIKE+J_M0_unconstrained_v1.pdf", width=8.5, height=11)
plot_BioGeoBEARS_results(resDIVALIKE, "BioGeoBEARS DIVALIKE on Psychotria M0_unconstrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDIVALIKE, "BioGeoBEARS DIVALIKE on Psychotria M0_unconstrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDIVALIKEj, "BioGeoBEARS DIVALIKE+J on Psychotria M0_unconstrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDIVALIKEj, "BioGeoBEARS DIVALIKE+J on Psychotria M0_unconstrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
dev.off()


# ------------------------------------------------------------------------------
# 4. BAYAREALIKE & BAYAREALIKE+J MODELS
# ------------------------------------------------------------------------------

# --- Run BAYAREALIKE ---
bgb_bay <- setup_base_bgb(trfn, geogfn, max_range_size)

# Set up BAYAREALIKE specific parameters (No subset sympatry, no vicariance)
bgb_bay$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_bay$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_bay$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_bay$BioGeoBEARS_model_object@params_table["v", "type"] <- "fixed"
bgb_bay$BioGeoBEARS_model_object@params_table["v", "init"] <- 0.0
bgb_bay$BioGeoBEARS_model_object@params_table["v", "est"] <- 0.0

# Adjust linkage
bgb_bay$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "1-j"
bgb_bay$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/1"
bgb_bay$BioGeoBEARS_model_object@params_table["y", "type"] <- "1-j"

# Only sympatric/range-copying events allowed
bgb_bay$BioGeoBEARS_model_object@params_table["mx01y", "type"] <- "fixed"
bgb_bay$BioGeoBEARS_model_object@params_table["mx01y", "init"] <- 0.9999
bgb_bay$BioGeoBEARS_model_object@params_table["mx01y", "est"] <- 0.9999

check_BioGeoBEARS_run(bgb_bay)

resfn_bay <- "Psychotria_BAYAREALIKE_M0_unconstrained_v1.Rdata"
if (!file.exists(resfn_bay)) {
  resBAYAREALIKE <- bears_optim_run(bgb_bay)
  save(resBAYAREALIKE, file = resfn_bay)
} else {
  load(resfn_bay)
}

# --- Run BAYAREALIKE+J ---
bgb_bayj <- setup_base_bgb(trfn, geogfn, max_range_size)

# Use BAYAREALIKE estimates as starting values
dstart <- resBAYAREALIKE$outputs@params_table["d","est"]
estart <- resBAYAREALIKE$outputs@params_table["e","est"]
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "init"] <- dstart
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "est"] <- dstart
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "init"] <- estart
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "est"] <- estart

# Set up BAYAREALIKE+J specific parameters
bgb_bayj$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_bayj$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_bayj$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_bayj$BioGeoBEARS_model_object@params_table["v", "type"] <- "fixed"
bgb_bayj$BioGeoBEARS_model_object@params_table["v", "init"] <- 0.0
bgb_bayj$BioGeoBEARS_model_object@params_table["v", "est"] <- 0.0

bgb_bayj$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "1-j"
bgb_bayj$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/1"
bgb_bayj$BioGeoBEARS_model_object@params_table["y", "type"] <- "1-j"

bgb_bayj$BioGeoBEARS_model_object@params_table["mx01y", "type"] <- "fixed"
bgb_bayj$BioGeoBEARS_model_object@params_table["mx01y", "init"] <- 0.9999
bgb_bayj$BioGeoBEARS_model_object@params_table["mx01y", "est"] <- 0.9999

# Add jump dispersal (j)
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "type"] <- "free"
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "init"] <- 0.0001
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "est"] <- 0.0001
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "max"] <- 0.99999

# Safe boundaries to prevent precision underruns in optim
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "min"] <- 0.0000001
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "max"] <- 4.9999999
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "min"] <- 0.0000001
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "max"] <- 4.9999999
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "min"] <- 0.00001

check_BioGeoBEARS_run(bgb_bayj)

resfn_bayj <- "Psychotria_BAYAREALIKE+J_M0_unconstrained_v1.Rdata"
if (!file.exists(resfn_bayj)) {
  resBAYAREALIKEj <- bears_optim_run(bgb_bayj)
  save(resBAYAREALIKEj, file = resfn_bayj)
} else {
  load(resfn_bayj)
}

# --- Plot BAYAREALIKE & BAYAREALIKE+J ---
pdf("Psychotria_BAYAREALIKE_vs_BAYAREALIKE+J_M0_unconstrained_v1.pdf", width=8.5, height=11)
plot_BioGeoBEARS_results(resBAYAREALIKE, "BioGeoBEARS BAYAREALIKE on Psychotria M0_unconstrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resBAYAREALIKE, "BioGeoBEARS BAYAREALIKE on Psychotria M0_unconstrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resBAYAREALIKEj, "BioGeoBEARS BAYAREALIKE+J on Psychotria M0_unconstrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resBAYAREALIKEj, "BioGeoBEARS BAYAREALIKE+J on Psychotria M0_unconstrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
dev.off()


# ------------------------------------------------------------------------------
# 5. MODEL COMPARISON (AIC & LIKELIHOOD RATIO TESTS)
# ------------------------------------------------------------------------------

# Setup empty tables
restable <- NULL
teststable <- NULL

# Extract parameters function wrapper for cleaner code
extract_params <- function(res_obj) {
  extract_params_from_BioGeoBEARS_results_object(results_object=res_obj, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
}

# DEC vs DEC+J
LnL_dec  <- get_LnL_from_BioGeoBEARS_results_object(resDEC)
LnL_decj <- get_LnL_from_BioGeoBEARS_results_object(resDECj)
stats_dec <- AICstats_2models(LnL_decj, LnL_dec, 3, 2)
restable <- rbind(restable, extract_params(resDEC), extract_params(resDECj))
teststable <- rbind(teststable, conditional_format_table(stats_dec))

# DIVALIKE vs DIVALIKE+J
LnL_diva  <- get_LnL_from_BioGeoBEARS_results_object(resDIVALIKE)
LnL_divaj <- get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEj)
stats_diva <- AICstats_2models(LnL_divaj, LnL_diva, 3, 2)
restable <- rbind(restable, extract_params(resDIVALIKE), extract_params(resDIVALIKEj))
teststable <- rbind(teststable, conditional_format_table(stats_diva))

# BAYAREALIKE vs BAYAREALIKE+J
LnL_bay  <- get_LnL_from_BioGeoBEARS_results_object(resBAYAREALIKE)
LnL_bayj <- get_LnL_from_BioGeoBEARS_results_object(resBAYAREALIKEj)
stats_bay <- AICstats_2models(LnL_bayj, LnL_bay, 3, 2)
restable <- rbind(restable, extract_params(resBAYAREALIKE), extract_params(resBAYAREALIKEj))
teststable <- rbind(teststable, conditional_format_table(stats_bay))

# Format and label final results tables
teststable$alt <- c("DEC+J", "DIVALIKE+J", "BAYAREALIKE+J")
teststable$null <- c("DEC", "DIVALIKE", "BAYAREALIKE")
row.names(restable) <- c("DEC", "DEC+J", "DIVALIKE", "DIVALIKE+J", "BAYAREALIKE", "BAYAREALIKE+J")

# Print final results
print(restable)
print(teststable)

# Save tables
save(restable, file="restable_M0_unconstrained_v1.Rdata")
save(teststable, file="teststable_M0_unconstrained_v1.Rdata")
```


Análisis Biogeográfico Restringido (M1_constrained)

En esta segunda fase, vamos a ejecutar exactamente los mismos modelos evolutivos (DEC, DIVALIKE, BAYAREALIKE y sus variantes con salto +J), pero introduciremos una restricción biológica/geográfica: limitaremos el tamaño máximo del rango a 2 áreas (max_range_size = 2). Llamaremos a esta configuración M1_constrained.

Al comparar los resultados estadísticos (AIC y LRT) de esta corrida con la anterior (M0_unconstrained), podremos evaluar si restringir el número de áreas ocupadas simultáneamente mejora el ajuste del modelo a nuestros datos de Psychotria.

```{r}

################################################################################
# BIOGEOBEARS ANALYSIS: M1_CONSTRAINED
# Models: DEC, DEC+J, DIVALIKE, DIVALIKE+J, BAYAREALIKE, BAYAREALIKE+J
# Constraint: max_range_size = 2
################################################################################

# ------------------------------------------------------------------------------
# 1. SETUP AND DATA LOADING
# ------------------------------------------------------------------------------

# Load required libraries
library(rexpokit)
library(cladoRcpp)
library(ape)
library(BioGeoBEARS)

# Set working directory (Update this path to your actual working directory)
wd <- normalizePath("./")
setwd(wd)

# Define file paths for the phylogeny and geography data
trfn   <- normalizePath("./Psychotria_5.2.newick")
geogfn <- normalizePath("./Psychotria_geog.data")

# Read and inspect the phylogeny and geographic range data
tr <- read.tree(trfn)
tipranges <- getranges_from_LagrangePHYLIP(lgdata_fn = geogfn)

# Set the maximum number of areas any species may occupy (CONSTRAINED TO 2)
max_range_size <- 2

# Location of BioGeoBEARS plot scripts
scriptdir <- normalizePath(system.file("extdata/a_scripts", package="BioGeoBEARS"))


# ------------------------------------------------------------------------------
# HELPER FUNCTION FOR CLEAN INITIALIZATION
# ------------------------------------------------------------------------------
# This function generates a fresh, clean BioGeoBEARS run object to avoid S4 class 
# inheritance errors when copying objects between models.
setup_base_bgb <- function(tree_file, geog_file, max_size) {
  bgb <- define_BioGeoBEARS_run()
  bgb$trfn <- tree_file
  bgb$geogfn <- geog_file
  bgb$max_range_size <- max_size
  bgb$speedup <- TRUE
  bgb$use_optimx <- TRUE
  bgb$num_cores_to_use <- 1
  bgb$force_sparse <- FALSE
  bgb$calc_ancprobs <- TRUE
  bgb$return_condlikes_table <- TRUE
  bgb$calc_TTL_loglike_from_condlikes_table <- TRUE
  bgb <- readfiles_BioGeoBEARS_run(bgb)
  return(bgb)
}


# ------------------------------------------------------------------------------
# 2. DEC & DEC+J MODELS
# ------------------------------------------------------------------------------

# --- Run DEC ---
bgb_dec <- setup_base_bgb(trfn, geogfn, max_range_size)
check_BioGeoBEARS_run(bgb_dec)

resfn_dec <- "Psychotria_DEC_M1_constrained_v1.Rdata"
if (!file.exists(resfn_dec)) {
  resDEC <- bears_optim_run(bgb_dec)
  save(resDEC, file = resfn_dec)
} else {
  load(resfn_dec)
}

# --- Run DEC+J ---
bgb_decj <- setup_base_bgb(trfn, geogfn, max_range_size) 

# Use DEC estimates as starting values for DEC+J
dstart <- resDEC$outputs@params_table["d","est"]
estart <- resDEC$outputs@params_table["e","est"]

bgb_decj$BioGeoBEARS_model_object@params_table["d", "init"] <- dstart
bgb_decj$BioGeoBEARS_model_object@params_table["d", "est"] <- dstart
bgb_decj$BioGeoBEARS_model_object@params_table["e", "init"] <- estart
bgb_decj$BioGeoBEARS_model_object@params_table["e", "est"] <- estart

# Add jump dispersal (j) as a free parameter
bgb_decj$BioGeoBEARS_model_object@params_table["j", "type"] <- "free"
bgb_decj$BioGeoBEARS_model_object@params_table["j", "init"] <- 0.0001
bgb_decj$BioGeoBEARS_model_object@params_table["j", "est"] <- 0.0001

check_BioGeoBEARS_run(bgb_decj)

resfn_decj <- "Psychotria_DEC+J_M1_constrained_v1.Rdata"
if (!file.exists(resfn_decj)) {
  resDECj <- bears_optim_run(bgb_decj)
  save(resDECj, file = resfn_decj)
} else {
  load(resfn_decj)
}

# --- Plot DEC & DEC+J ---
pdf("Psychotria_DEC_vs_DEC+J_M1_constrained_v1.pdf", width=8.5, height=11)
plot_BioGeoBEARS_results(resDEC, "BioGeoBEARS DEC on Psychotria M1_constrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDEC, "BioGeoBEARS DEC on Psychotria M1_constrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDECj, "BioGeoBEARS DEC+J on Psychotria M1_constrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDECj, "BioGeoBEARS DEC+J on Psychotria M1_constrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
dev.off()


# ------------------------------------------------------------------------------
# 3. DIVALIKE & DIVALIKE+J MODELS
# ------------------------------------------------------------------------------

# --- Run DIVALIKE ---
bgb_diva <- setup_base_bgb(trfn, geogfn, max_range_size) 

# Set up DIVALIKE specific parameters (Remove subset-sympatry)
bgb_diva$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_diva$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_diva$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_diva$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "2-j"
bgb_diva$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/2"
bgb_diva$BioGeoBEARS_model_object@params_table["y", "type"] <- "ysv*1/2"
bgb_diva$BioGeoBEARS_model_object@params_table["v", "type"] <- "ysv*1/2"

# Allow classic widespread vicariance
bgb_diva$BioGeoBEARS_model_object@params_table["mx01v", "type"] <- "fixed"
bgb_diva$BioGeoBEARS_model_object@params_table["mx01v", "init"] <- 0.5
bgb_diva$BioGeoBEARS_model_object@params_table["mx01v", "est"] <- 0.5

check_BioGeoBEARS_run(bgb_diva)

resfn_diva <- "Psychotria_DIVALIKE_M1_constrained_v1.Rdata"
if (!file.exists(resfn_diva)) {
  resDIVALIKE <- bears_optim_run(bgb_diva)
  save(resDIVALIKE, file = resfn_diva)
} else {
  load(resfn_diva)
}

# --- Run DIVALIKE+J ---
bgb_divaj <- setup_base_bgb(trfn, geogfn, max_range_size) 

# Use DIVALIKE estimates as starting values
dstart <- resDIVALIKE$outputs@params_table["d","est"]
estart <- resDIVALIKE$outputs@params_table["e","est"]
bgb_divaj$BioGeoBEARS_model_object@params_table["d", "init"] <- dstart
bgb_divaj$BioGeoBEARS_model_object@params_table["d", "est"] <- dstart
bgb_divaj$BioGeoBEARS_model_object@params_table["e", "init"] <- estart
bgb_divaj$BioGeoBEARS_model_object@params_table["e", "est"] <- estart

# Set up DIVALIKE+J specific parameters
bgb_divaj$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_divaj$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_divaj$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_divaj$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "2-j"
bgb_divaj$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/2"
bgb_divaj$BioGeoBEARS_model_object@params_table["y", "type"] <- "ysv*1/2"
bgb_divaj$BioGeoBEARS_model_object@params_table["v", "type"] <- "ysv*1/2"

bgb_divaj$BioGeoBEARS_model_object@params_table["mx01v", "type"] <- "fixed"
bgb_divaj$BioGeoBEARS_model_object@params_table["mx01v", "init"] <- 0.5
bgb_divaj$BioGeoBEARS_model_object@params_table["mx01v", "est"] <- 0.5

# Add jump dispersal (j)
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "type"] <- "free"
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "init"] <- 0.0001
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "est"] <- 0.0001
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "min"] <- 0.00001
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "max"] <- 1.99999

check_BioGeoBEARS_run(bgb_divaj)

resfn_divaj <- "Psychotria_DIVALIKE+J_M1_constrained_v1.Rdata"
if (!file.exists(resfn_divaj)) {
  resDIVALIKEj <- bears_optim_run(bgb_divaj)
  save(resDIVALIKEj, file = resfn_divaj)
} else {
  load(resfn_divaj)
}

# --- Plot DIVALIKE & DIVALIKE+J ---
pdf("Psychotria_DIVALIKE_vs_DIVALIKE+J_M1_constrained_v1.pdf", width=8.5, height=11)
plot_BioGeoBEARS_results(resDIVALIKE, "BioGeoBEARS DIVALIKE on Psychotria M1_constrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDIVALIKE, "BioGeoBEARS DIVALIKE on Psychotria M1_constrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDIVALIKEj, "BioGeoBEARS DIVALIKE+J on Psychotria M1_constrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDIVALIKEj, "BioGeoBEARS DIVALIKE+J on Psychotria M1_constrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
dev.off()


# ------------------------------------------------------------------------------
# 4. BAYAREALIKE & BAYAREALIKE+J MODELS
# ------------------------------------------------------------------------------

# --- Run BAYAREALIKE ---
bgb_bay <- setup_base_bgb(trfn, geogfn, max_range_size)

# Set up BAYAREALIKE specific parameters (No subset sympatry, no vicariance)
bgb_bay$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_bay$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_bay$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_bay$BioGeoBEARS_model_object@params_table["v", "type"] <- "fixed"
bgb_bay$BioGeoBEARS_model_object@params_table["v", "init"] <- 0.0
bgb_bay$BioGeoBEARS_model_object@params_table["v", "est"] <- 0.0

# Adjust linkage
bgb_bay$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "1-j"
bgb_bay$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/1"
bgb_bay$BioGeoBEARS_model_object@params_table["y", "type"] <- "1-j"

# Only sympatric/range-copying events allowed
bgb_bay$BioGeoBEARS_model_object@params_table["mx01y", "type"] <- "fixed"
bgb_bay$BioGeoBEARS_model_object@params_table["mx01y", "init"] <- 0.9999
bgb_bay$BioGeoBEARS_model_object@params_table["mx01y", "est"] <- 0.9999

check_BioGeoBEARS_run(bgb_bay)

resfn_bay <- "Psychotria_BAYAREALIKE_M1_constrained_v1.Rdata"
if (!file.exists(resfn_bay)) {
  resBAYAREALIKE <- bears_optim_run(bgb_bay)
  save(resBAYAREALIKE, file = resfn_bay)
} else {
  load(resfn_bay)
}

# --- Run BAYAREALIKE+J ---
bgb_bayj <- setup_base_bgb(trfn, geogfn, max_range_size)

# Use BAYAREALIKE estimates as starting values
dstart <- resBAYAREALIKE$outputs@params_table["d","est"]
estart <- resBAYAREALIKE$outputs@params_table["e","est"]
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "init"] <- dstart
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "est"] <- dstart
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "init"] <- estart
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "est"] <- estart

# Set up BAYAREALIKE+J specific parameters
bgb_bayj$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_bayj$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_bayj$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_bayj$BioGeoBEARS_model_object@params_table["v", "type"] <- "fixed"
bgb_bayj$BioGeoBEARS_model_object@params_table["v", "init"] <- 0.0
bgb_bayj$BioGeoBEARS_model_object@params_table["v", "est"] <- 0.0

bgb_bayj$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "1-j"
bgb_bayj$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/1"
bgb_bayj$BioGeoBEARS_model_object@params_table["y", "type"] <- "1-j"

bgb_bayj$BioGeoBEARS_model_object@params_table["mx01y", "type"] <- "fixed"
bgb_bayj$BioGeoBEARS_model_object@params_table["mx01y", "init"] <- 0.9999
bgb_bayj$BioGeoBEARS_model_object@params_table["mx01y", "est"] <- 0.9999

# Add jump dispersal (j)
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "type"] <- "free"
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "init"] <- 0.0001
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "est"] <- 0.0001
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "max"] <- 0.99999

# Safe boundaries to prevent precision underruns in optim
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "min"] <- 0.0000001
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "max"] <- 4.9999999
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "min"] <- 0.0000001
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "max"] <- 4.9999999
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "min"] <- 0.00001

check_BioGeoBEARS_run(bgb_bayj)

resfn_bayj <- "Psychotria_BAYAREALIKE+J_M1_constrained_v1.Rdata"
if (!file.exists(resfn_bayj)) {
  resBAYAREALIKEj <- bears_optim_run(bgb_bayj)
  save(resBAYAREALIKEj, file = resfn_bayj)
} else {
  load(resfn_bayj)
}

# --- Plot BAYAREALIKE & BAYAREALIKE+J ---
pdf("Psychotria_BAYAREALIKE_vs_BAYAREALIKE+J_M1_constrained_v1.pdf", width=8.5, height=11)
plot_BioGeoBEARS_results(resBAYAREALIKE, "BioGeoBEARS BAYAREALIKE on Psychotria M1_constrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resBAYAREALIKE, "BioGeoBEARS BAYAREALIKE on Psychotria M1_constrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resBAYAREALIKEj, "BioGeoBEARS BAYAREALIKE+J on Psychotria M1_constrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resBAYAREALIKEj, "BioGeoBEARS BAYAREALIKE+J on Psychotria M1_constrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
dev.off()


# ------------------------------------------------------------------------------
# 5. MODEL COMPARISON (AIC & LIKELIHOOD RATIO TESTS)
# ------------------------------------------------------------------------------

# Setup empty tables
restable <- NULL
teststable <- NULL

# Extract parameters function wrapper for cleaner code
extract_params <- function(res_obj) {
  extract_params_from_BioGeoBEARS_results_object(results_object=res_obj, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
}

# DEC vs DEC+J
LnL_dec  <- get_LnL_from_BioGeoBEARS_results_object(resDEC)
LnL_decj <- get_LnL_from_BioGeoBEARS_results_object(resDECj)
stats_dec <- AICstats_2models(LnL_decj, LnL_dec, 3, 2)
restable <- rbind(restable, extract_params(resDEC), extract_params(resDECj))
teststable <- rbind(teststable, conditional_format_table(stats_dec))

# DIVALIKE vs DIVALIKE+J
LnL_diva  <- get_LnL_from_BioGeoBEARS_results_object(resDIVALIKE)
LnL_divaj <- get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEj)
stats_diva <- AICstats_2models(LnL_divaj, LnL_diva, 3, 2)
restable <- rbind(restable, extract_params(resDIVALIKE), extract_params(resDIVALIKEj))
teststable <- rbind(teststable, conditional_format_table(stats_diva))

# BAYAREALIKE vs BAYAREALIKE+J
LnL_bay  <- get_LnL_from_BioGeoBEARS_results_object(resBAYAREALIKE)
LnL_bayj <- get_LnL_from_BioGeoBEARS_results_object(resBAYAREALIKEj)
stats_bay <- AICstats_2models(LnL_bayj, LnL_bay, 3, 2)
restable <- rbind(restable, extract_params(resBAYAREALIKE), extract_params(resBAYAREALIKEj))
teststable <- rbind(teststable, conditional_format_table(stats_bay))

# Format and label final results tables
teststable$alt <- c("DEC+J", "DIVALIKE+J", "BAYAREALIKE+J")
teststable$null <- c("DEC", "DIVALIKE", "BAYAREALIKE")
row.names(restable) <- c("DEC", "DEC+J", "DIVALIKE", "DIVALIKE+J", "BAYAREALIKE", "BAYAREALIKE+J")

# Print final results
print(restable)
print(teststable)

# Save tables
save(restable, file="restable_M1_constrained_v1.Rdata")
save(teststable, file="teststable_M1_constrained_v1.Rdata")

```


Análisis Biogeográfico con Restricción Espacial (M2_constrained - Stepping Stones)

En esta tercera iteración, ejecutaremos nuevamente la batería de modelos (DEC, DIVALIKE, BAYAREALIKE y sus variantes +J), pero esta vez bajo un modelo de "stepping stones" (islas puente).

En este escenario, aplicaremos dos restricciones fundamentales para reflejar la geología del archipiélago de Hawái:

    Restricción de tamaño de rango: Limitaremos el tamaño máximo del rango a 2 áreas (max_range_size = 2), igual que en el modelo M1.

    Restricción direccional de dispersión: Proveeremos una matriz de multiplicadores de dispersión (Psychotria-steppingstone-dispersal.txt). Esta matriz establece que las transiciones solo son posibles entre islas geográficamente adyacentes y, además, siguen una direccionalidad estricta desde las islas más antiguas hacia las más nuevas (de este a oeste).

Llamaremos a esta configuración M2_constrained

```{r}

################################################################################
# BIOGEOBEARS ANALYSIS: M2_CONSTRAINED (STEPPING STONES)
# Models: DEC, DEC+J, DIVALIKE, DIVALIKE+J, BAYAREALIKE, BAYAREALIKE+J
# Constraints: max_range_size = 2 AND stepping-stone dispersal matrix
################################################################################

# ------------------------------------------------------------------------------
# 1. SETUP AND DATA LOADING
# ------------------------------------------------------------------------------

# Load required libraries
library(rexpokit)
library(cladoRcpp)
library(ape)
library(BioGeoBEARS)

# Set working directory (Update this path to your actual working directory)
wd <- normalizePath("./")
setwd(wd)

# Define file paths for the phylogeny, geography data, and dispersal matrix
trfn   <- normalizePath("./Psychotria_5.2.newick")
geogfn <- normalizePath("./Psychotria_geog.data")
disp_matrix_fn <- "Psychotria-steppingstone-dispersal.txt" # Make sure this file is in your working directory

# Read and inspect the phylogeny and geographic range data
tr <- read.tree(trfn)
tipranges <- getranges_from_LagrangePHYLIP(lgdata_fn = geogfn)

# Set the maximum number of areas any species may occupy (CONSTRAINED TO 2)
max_range_size <- 2

# INTERRUPTOR CLAVE: Forzar recálculo para ignorar la caché de errores previos
force_recalculate <- TRUE 

# Location of BioGeoBEARS plot scripts
scriptdir <- normalizePath(system.file("extdata/a_scripts", package="BioGeoBEARS"))


# ------------------------------------------------------------------------------
# HELPER FUNCTION FOR CLEAN INITIALIZATION WITH DISPERSAL MATRIX
# ------------------------------------------------------------------------------
# This function generates a fresh, clean BioGeoBEARS run object including the 
# stepping-stone dispersal matrix.
# FUNCIÓN CORREGIDA PARA M2
setup_base_bgb_M2 <- function(tree_file, geog_file, max_size, disp_file) {
  bgb <- define_BioGeoBEARS_run()
  bgb$trfn <- tree_file
  bgb$geogfn <- geog_file
  bgb$max_range_size <- max_size
  
  # Asignamos el archivo ANTES de leer
  bgb$dispersal_multipliers_fn <- disp_file 
  
  bgb$speedup <- TRUE
  bgb$use_optimx <- TRUE
  bgb$num_cores_to_use <- 1
  bgb$force_sparse <- FALSE
  bgb$calc_ancprobs <- TRUE
  bgb$return_condlikes_table <- TRUE
  bgb$calc_TTL_loglike_from_condlikes_table <- TRUE
  
  # AHORA LEEMOS LOS ARCHIVOS (incluyendo la matriz de dispersión)
  bgb <- readfiles_BioGeoBEARS_run(bgb)
  return(bgb)
}


# ------------------------------------------------------------------------------
# 2. DEC & DEC+J MODELS
# ------------------------------------------------------------------------------

# --- Run DEC ---
bgb_dec <- setup_base_bgb_M2(trfn, geogfn, max_range_size, disp_matrix_fn)
check_BioGeoBEARS_run(bgb_dec)

resfn_dec <- "Psychotria_DEC_M2_constrained_v1.Rdata"
if (!file.exists(resfn_dec) | force_recalculate) {
  resDEC <- bears_optim_run(bgb_dec)
  save(resDEC, file = resfn_dec)
} else {
  load(resfn_dec)
}

# --- Run DEC+J ---
bgb_decj <- setup_base_bgb_M2(trfn, geogfn, max_range_size, disp_matrix_fn) 

# Use DEC estimates as starting values for DEC+J
dstart <- resDEC$outputs@params_table["d","est"]
estart <- resDEC$outputs@params_table["e","est"]

bgb_decj$BioGeoBEARS_model_object@params_table["d", "init"] <- dstart
bgb_decj$BioGeoBEARS_model_object@params_table["d", "est"] <- dstart
bgb_decj$BioGeoBEARS_model_object@params_table["e", "init"] <- estart
bgb_decj$BioGeoBEARS_model_object@params_table["e", "est"] <- estart

# Add jump dispersal (j) as a free parameter
bgb_decj$BioGeoBEARS_model_object@params_table["j", "type"] <- "free"
bgb_decj$BioGeoBEARS_model_object@params_table["j", "init"] <- 0.0001
bgb_decj$BioGeoBEARS_model_object@params_table["j", "est"] <- 0.0001

check_BioGeoBEARS_run(bgb_decj)

resfn_decj <- "Psychotria_DEC+J_M2_constrained_v1.Rdata"
if (!file.exists(resfn_decj) | force_recalculate) {
  resDECj <- bears_optim_run(bgb_decj)
  save(resDECj, file = resfn_decj)
} else {
  load(resfn_decj)
}

# --- Plot DEC & DEC+J ---
pdf("Psychotria_DEC_vs_DEC+J_M2_constrained_v1.pdf", width=8.5, height=11)
plot_BioGeoBEARS_results(resDEC, "BioGeoBEARS DEC on Psychotria M2_constrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDEC, "BioGeoBEARS DEC on Psychotria M2_constrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDECj, "BioGeoBEARS DEC+J on Psychotria M2_constrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDECj, "BioGeoBEARS DEC+J on Psychotria M2_constrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
dev.off()


# ------------------------------------------------------------------------------
# 3. DIVALIKE & DIVALIKE+J MODELS
# ------------------------------------------------------------------------------

# --- Run DIVALIKE ---
bgb_diva <- setup_base_bgb_M2(trfn, geogfn, max_range_size, disp_matrix_fn) 

# Set up DIVALIKE specific parameters (Remove subset-sympatry)
bgb_diva$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_diva$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_diva$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_diva$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "2-j"
bgb_diva$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/2"
bgb_diva$BioGeoBEARS_model_object@params_table["y", "type"] <- "ysv*1/2"
bgb_diva$BioGeoBEARS_model_object@params_table["v", "type"] <- "ysv*1/2"

# Allow classic widespread vicariance
bgb_diva$BioGeoBEARS_model_object@params_table["mx01v", "type"] <- "fixed"
bgb_diva$BioGeoBEARS_model_object@params_table["mx01v", "init"] <- 0.5
bgb_diva$BioGeoBEARS_model_object@params_table["mx01v", "est"] <- 0.5

check_BioGeoBEARS_run(bgb_diva)

resfn_diva <- "Psychotria_DIVALIKE_M2_constrained_v1.Rdata"
if (!file.exists(resfn_diva) | force_recalculate) {
  resDIVALIKE <- bears_optim_run(bgb_diva)
  save(resDIVALIKE, file = resfn_diva)
} else {
  load(resfn_diva)
}

# --- Run DIVALIKE+J ---
bgb_divaj <- setup_base_bgb_M2(trfn, geogfn, max_range_size, disp_matrix_fn) 

# Use DIVALIKE estimates as starting values
dstart <- resDIVALIKE$outputs@params_table["d","est"]
estart <- resDIVALIKE$outputs@params_table["e","est"]
bgb_divaj$BioGeoBEARS_model_object@params_table["d", "init"] <- dstart
bgb_divaj$BioGeoBEARS_model_object@params_table["d", "est"] <- dstart
bgb_divaj$BioGeoBEARS_model_object@params_table["e", "init"] <- estart
bgb_divaj$BioGeoBEARS_model_object@params_table["e", "est"] <- estart

# Set up DIVALIKE+J specific parameters
bgb_divaj$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_divaj$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_divaj$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_divaj$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "2-j"
bgb_divaj$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/2"
bgb_divaj$BioGeoBEARS_model_object@params_table["y", "type"] <- "ysv*1/2"
bgb_divaj$BioGeoBEARS_model_object@params_table["v", "type"] <- "ysv*1/2"

bgb_divaj$BioGeoBEARS_model_object@params_table["mx01v", "type"] <- "fixed"
bgb_divaj$BioGeoBEARS_model_object@params_table["mx01v", "init"] <- 0.5
bgb_divaj$BioGeoBEARS_model_object@params_table["mx01v", "est"] <- 0.5

# Add jump dispersal (j)
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "type"] <- "free"
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "init"] <- 0.0001
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "est"] <- 0.0001
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "min"] <- 0.00001
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "max"] <- 1.99999

check_BioGeoBEARS_run(bgb_divaj)

resfn_divaj <- "Psychotria_DIVALIKE+J_M2_constrained_v1.Rdata"
if (!file.exists(resfn_divaj) | force_recalculate) {
  resDIVALIKEj <- bears_optim_run(bgb_divaj)
  save(resDIVALIKEj, file = resfn_divaj)
} else {
  load(resfn_divaj)
}

# --- Plot DIVALIKE & DIVALIKE+J ---
pdf("Psychotria_DIVALIKE_vs_DIVALIKE+J_M2_constrained_v1.pdf", width=8.5, height=11)
plot_BioGeoBEARS_results(resDIVALIKE, "BioGeoBEARS DIVALIKE on Psychotria M2_constrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDIVALIKE, "BioGeoBEARS DIVALIKE on Psychotria M2_constrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDIVALIKEj, "BioGeoBEARS DIVALIKE+J on Psychotria M2_constrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDIVALIKEj, "BioGeoBEARS DIVALIKE+J on Psychotria M2_constrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
dev.off()


# ------------------------------------------------------------------------------
# 4. BAYAREALIKE & BAYAREALIKE+J MODELS
# ------------------------------------------------------------------------------

# --- Run BAYAREALIKE ---
bgb_bay <- setup_base_bgb_M2(trfn, geogfn, max_range_size, disp_matrix_fn)

# Set up BAYAREALIKE specific parameters (No subset sympatry, no vicariance)
bgb_bay$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_bay$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_bay$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_bay$BioGeoBEARS_model_object@params_table["v", "type"] <- "fixed"
bgb_bay$BioGeoBEARS_model_object@params_table["v", "init"] <- 0.0
bgb_bay$BioGeoBEARS_model_object@params_table["v", "est"] <- 0.0

# Adjust linkage
bgb_bay$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "1-j"
bgb_bay$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/1"
bgb_bay$BioGeoBEARS_model_object@params_table["y", "type"] <- "1-j"

# Only sympatric/range-copying events allowed
bgb_bay$BioGeoBEARS_model_object@params_table["mx01y", "type"] <- "fixed"
bgb_bay$BioGeoBEARS_model_object@params_table["mx01y", "init"] <- 0.9999
bgb_bay$BioGeoBEARS_model_object@params_table["mx01y", "est"] <- 0.9999

check_BioGeoBEARS_run(bgb_bay)

resfn_bay <- "Psychotria_BAYAREALIKE_M2_constrained_v1.Rdata"
if (!file.exists(resfn_bay) | force_recalculate) {
  resBAYAREALIKE <- bears_optim_run(bgb_bay)
  save(resBAYAREALIKE, file = resfn_bay)
} else {
  load(resfn_bay)
}

# --- Run BAYAREALIKE+J ---
bgb_bayj <- setup_base_bgb_M2(trfn, geogfn, max_range_size, disp_matrix_fn)

# Use BAYAREALIKE estimates as starting values
dstart <- resBAYAREALIKE$outputs@params_table["d","est"]
estart <- resBAYAREALIKE$outputs@params_table["e","est"]
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "init"] <- dstart
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "est"] <- dstart
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "init"] <- estart
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "est"] <- estart

# Set up BAYAREALIKE+J specific parameters
bgb_bayj$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_bayj$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_bayj$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_bayj$BioGeoBEARS_model_object@params_table["v", "type"] <- "fixed"
bgb_bayj$BioGeoBEARS_model_object@params_table["v", "init"] <- 0.0
bgb_bayj$BioGeoBEARS_model_object@params_table["v", "est"] <- 0.0

bgb_bayj$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "1-j"
bgb_bayj$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/1"
bgb_bayj$BioGeoBEARS_model_object@params_table["y", "type"] <- "1-j"

bgb_bayj$BioGeoBEARS_model_object@params_table["mx01y", "type"] <- "fixed"
bgb_bayj$BioGeoBEARS_model_object@params_table["mx01y", "init"] <- 0.9999
bgb_bayj$BioGeoBEARS_model_object@params_table["mx01y", "est"] <- 0.9999

# Add jump dispersal (j)
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "type"] <- "free"
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "init"] <- 0.0001
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "est"] <- 0.0001
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "max"] <- 0.99999

# Safe boundaries to prevent precision underruns in optim
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "min"] <- 0.0000001
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "max"] <- 4.9999999
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "min"] <- 0.0000001
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "max"] <- 4.9999999
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "min"] <- 0.00001

check_BioGeoBEARS_run(bgb_bayj)

resfn_bayj <- "Psychotria_BAYAREALIKE+J_M2_constrained_v1.Rdata"
if (!file.exists(resfn_bayj) | force_recalculate) {
  resBAYAREALIKEj <- bears_optim_run(bgb_bayj)
  save(resBAYAREALIKEj, file = resfn_bayj)
} else {
  load(resfn_bayj)
}

# --- Plot BAYAREALIKE & BAYAREALIKE+J ---
pdf("Psychotria_BAYAREALIKE_vs_BAYAREALIKE+J_M2_constrained_v1.pdf", width=8.5, height=11)
plot_BioGeoBEARS_results(resBAYAREALIKE, "BioGeoBEARS BAYAREALIKE on Psychotria M2_constrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resBAYAREALIKE, "BioGeoBEARS BAYAREALIKE on Psychotria M2_constrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resBAYAREALIKEj, "BioGeoBEARS BAYAREALIKE+J on Psychotria M2_constrained", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resBAYAREALIKEj, "BioGeoBEARS BAYAREALIKE+J on Psychotria M2_constrained", addl_params=list("j"), plotwhat="pie", cornercoords_loc=scriptdir, tr=tr, tipranges=tipranges)
dev.off()


# ------------------------------------------------------------------------------
# 5. MODEL COMPARISON (AIC & LIKELIHOOD RATIO TESTS)
# ------------------------------------------------------------------------------

# Setup empty tables
restable <- NULL
teststable <- NULL

# Extract parameters function wrapper for cleaner code
extract_params <- function(res_obj) {
  extract_params_from_BioGeoBEARS_results_object(results_object=res_obj, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
}

# DEC vs DEC+J
LnL_dec  <- get_LnL_from_BioGeoBEARS_results_object(resDEC)
LnL_decj <- get_LnL_from_BioGeoBEARS_results_object(resDECj)
stats_dec <- AICstats_2models(LnL_decj, LnL_dec, 3, 2)
restable <- rbind(restable, extract_params(resDEC), extract_params(resDECj))
teststable <- rbind(teststable, conditional_format_table(stats_dec))

# DIVALIKE vs DIVALIKE+J
LnL_diva  <- get_LnL_from_BioGeoBEARS_results_object(resDIVALIKE)
LnL_divaj <- get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEj)
stats_diva <- AICstats_2models(LnL_divaj, LnL_diva, 3, 2)
restable <- rbind(restable, extract_params(resDIVALIKE), extract_params(resDIVALIKEj))
teststable <- rbind(teststable, conditional_format_table(stats_diva))

# BAYAREALIKE vs BAYAREALIKE+J
LnL_bay  <- get_LnL_from_BioGeoBEARS_results_object(resBAYAREALIKE)
LnL_bayj <- get_LnL_from_BioGeoBEARS_results_object(resBAYAREALIKEj)
stats_bay <- AICstats_2models(LnL_bayj, LnL_bay, 3, 2)
restable <- rbind(restable, extract_params(resBAYAREALIKE), extract_params(resBAYAREALIKEj))
teststable <- rbind(teststable, conditional_format_table(stats_bay))

# Format and label final results tables
teststable$alt <- c("DEC+J", "DIVALIKE+J", "BAYAREALIKE+J")
teststable$null <- c("DEC", "DIVALIKE", "BAYAREALIKE")
row.names(restable) <- c("DEC", "DEC+J", "DIVALIKE", "DIVALIKE+J", "BAYAREALIKE", "BAYAREALIKE+J")

# Print final results
print(restable)
print(teststable)

# Save tables
save(restable, file="restable_M2_constrained_v1.Rdata")
save(teststable, file="teststable_M2_constrained_v1.Rdata")

```

Análisis Biogeográfico Estratificado en el Tiempo y Mapeo Estocástico

En esta última etapa, vamos a implementar un modelo estratificado en el tiempo (M3_stratified). Esto significa que las dinámicas del modelo cambiarán en diferentes periodos temporales de la filogenia. Específicamente, restringiremos el modelo para que la dispersión a ciertas islas solo sea posible después de que estas hayan emergido geológicamente.

Existen varias formas de introducir estas restricciones biológicas y geológicas en BioGeoBEARS:

    Matriz de áreas permitidas (areas_allowed_fn): Permite o prohíbe la presencia en ciertas áreas dependiendo del estrato temporal (la que usaremos aquí).

    Matriz de adyacencia: Indica las probabilidades de dispersión entre áreas (valores de 0 a 1). Como esto puede ser subjetivo, se puede estimar un parámetro w que ajusta dicha matriz durante el análisis.

    Matriz de distancias: Modela la probabilidad de dispersión en función de la distancia física. También se puede ajustar estimando un parámetro x que modifica el efecto de estas distancias.

(Para más detalles sobre estos modelos avanzados, puedes consultar la wiki oficial: http://phylo.wikidot.com/biogeobears).


```{r}

# Set the maximum number of areas any species may occupy
max_range_size <- 2

# INTERRUPTOR CLAVE: Forzar recálculo para ignorar la caché de errores previos
force_recalculate <- TRUE 

# ------------------------------------------------------------------------------
# FUNCIÓN AUXILIAR PARA INICIALIZACIÓN ESTRATIFICADA
# ------------------------------------------------------------------------------
# FUNCIÓN CORREGIDA PARA M3
setup_base_bgb_M3 <- function(tree_file, geog_file, max_size) {
  bgb <- define_BioGeoBEARS_run()
  bgb$trfn <- tree_file
  bgb$geogfn <- geog_file
  bgb$max_range_size <- max_size
  
  # Asignamos TODOS los archivos de estratificación ANTES de leer
  bgb$timesfn <- "timeperiods.txt"
  bgb$areas_allowed_fn <- "Psychotria-stratified-dispersal.txt"
  # NOTA: En modelos estratificados, BioGeoBEARS suele requerir que definamos
  # la matriz de adyacencia o distancias para cada estrato, o al menos
  # que la matriz base sea explícita.
  
  bgb$speedup <- TRUE
  bgb$use_optimx <- TRUE
  bgb$num_cores_to_use <- 1
  bgb$force_sparse <- FALSE
  bgb$calc_ancprobs <- TRUE
  bgb$return_condlikes_table <- TRUE
  bgb$calc_TTL_loglike_from_condlikes_table <- TRUE
  
  # AHORA LEEMOS LOS ARCHIVOS
  bgb <- readfiles_BioGeoBEARS_run(bgb)
  
  # SECCIONAMOS EL ÁRBOL (Requisito estricto de M3)
  bgb <- section_the_tree(inputs=bgb, make_master_table=TRUE, plot_pieces=FALSE)
  
  return(bgb)
}


# Define script directory for plots
scriptdir <- normalizePath(system.file("extdata/a_scripts", package="BioGeoBEARS"))

#######################################################
# DEC AND DEC+J ANALYSIS
#######################################################

# --- Run DEC ---
bgb_dec <- setup_base_bgb_M3(trfn, geogfn, max_range_size)
check_BioGeoBEARS_run(bgb_dec)

resfn_dec <- "Psychotria_DEC_M3_constrained_v1.Rdata"
if (!file.exists(resfn_dec) | force_recalculate) {
  resDEC <- bears_optim_run(bgb_dec)
  save(resDEC, file=resfn_dec)
} else {
  load(resfn_dec)
}

# --- Run DEC+J ---
bgb_decj <- setup_base_bgb_M3(trfn, geogfn, max_range_size)

dstart <- resDEC$outputs@params_table["d","est"]
estart <- resDEC$outputs@params_table["e","est"]

bgb_decj$BioGeoBEARS_model_object@params_table["d", "init"] <- dstart
bgb_decj$BioGeoBEARS_model_object@params_table["d", "est"] <- dstart
bgb_decj$BioGeoBEARS_model_object@params_table["e", "init"] <- estart
bgb_decj$BioGeoBEARS_model_object@params_table["e", "est"] <- estart

bgb_decj$BioGeoBEARS_model_object@params_table["j", "type"] <- "free"
bgb_decj$BioGeoBEARS_model_object@params_table["j", "init"] <- 0.0001
bgb_decj$BioGeoBEARS_model_object@params_table["j", "est"] <- 0.0001

check_BioGeoBEARS_run(bgb_decj)

resfn_decj <- "Psychotria_DEC+J_M3_constrained_v1.Rdata"
if (!file.exists(resfn_decj) | force_recalculate) {
  resDECj <- bears_optim_run(bgb_decj)
  save(resDECj, file=resfn_decj)
} else {
  load(resfn_decj)
}

pdf("Psychotria_DEC_vs_DEC+J_M3_constrained_v1.pdf", width=8.5, height=11)
analysis_titletxt ="BioGeoBEARS DEC on Psychotria M3_constrained"
plot_BioGeoBEARS_results(resDEC, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDEC, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

analysis_titletxt ="BioGeoBEARS DEC+J on Psychotria M3_constrained"
plot_BioGeoBEARS_results(resDECj, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDECj, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)
dev.off()


#######################################################
# DIVALIKE AND DIVALIKE+J ANALYSIS
#######################################################

# --- Run DIVALIKE ---
bgb_diva <- setup_base_bgb_M3(trfn, geogfn, max_range_size)

bgb_diva$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_diva$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_diva$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_diva$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "2-j"
bgb_diva$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/2"
bgb_diva$BioGeoBEARS_model_object@params_table["y", "type"] <- "ysv*1/2"
bgb_diva$BioGeoBEARS_model_object@params_table["v", "type"] <- "ysv*1/2"

bgb_diva$BioGeoBEARS_model_object@params_table["mx01v", "type"] <- "fixed"
bgb_diva$BioGeoBEARS_model_object@params_table["mx01v", "init"] <- 0.5
bgb_diva$BioGeoBEARS_model_object@params_table["mx01v", "est"] <- 0.5

check_BioGeoBEARS_run(bgb_diva)

resfn_diva <- "Psychotria_DIVALIKE_M3_constrained_v1.Rdata"
if (!file.exists(resfn_diva) | force_recalculate) {
  resDIVALIKE <- bears_optim_run(bgb_diva)
  save(resDIVALIKE, file=resfn_diva)
} else {
  load(resfn_diva)
}

# --- Run DIVALIKE+J ---
bgb_divaj <- setup_base_bgb_M3(trfn, geogfn, max_range_size)

dstart <- resDIVALIKE$outputs@params_table["d","est"]
estart <- resDIVALIKE$outputs@params_table["e","est"]

bgb_divaj$BioGeoBEARS_model_object@params_table["d", "init"] <- dstart
bgb_divaj$BioGeoBEARS_model_object@params_table["d", "est"] <- dstart
bgb_divaj$BioGeoBEARS_model_object@params_table["e", "init"] <- estart
bgb_divaj$BioGeoBEARS_model_object@params_table["e", "est"] <- estart

bgb_divaj$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_divaj$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_divaj$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_divaj$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "2-j"
bgb_divaj$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/2"
bgb_divaj$BioGeoBEARS_model_object@params_table["y", "type"] <- "ysv*1/2"
bgb_divaj$BioGeoBEARS_model_object@params_table["v", "type"] <- "ysv*1/2"

bgb_divaj$BioGeoBEARS_model_object@params_table["mx01v", "type"] <- "fixed"
bgb_divaj$BioGeoBEARS_model_object@params_table["mx01v", "init"] <- 0.5
bgb_divaj$BioGeoBEARS_model_object@params_table["mx01v", "est"] <- 0.5

bgb_divaj$BioGeoBEARS_model_object@params_table["j", "type"] <- "free"
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "init"] <- 0.0001
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "est"] <- 0.0001
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "min"] <- 0.00001
bgb_divaj$BioGeoBEARS_model_object@params_table["j", "max"] <- 1.99999

check_BioGeoBEARS_run(bgb_divaj)

resfn_divaj <- "Psychotria_DIVALIKE+J_M3_constrained_v1.Rdata"
if (!file.exists(resfn_divaj) | force_recalculate) {
  resDIVALIKEj <- bears_optim_run(bgb_divaj)
  save(resDIVALIKEj, file=resfn_divaj)
} else {
  load(resfn_divaj)
}

pdf("Psychotria_DIVALIKE_vs_DIVALIKE+J_M3_constrained_v1.pdf", width=8.5, height=11)
analysis_titletxt ="BioGeoBEARS DIVALIKE on Psychotria M3_constrained"
plot_BioGeoBEARS_results(resDIVALIKE, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDIVALIKE, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

analysis_titletxt ="BioGeoBEARS DIVALIKE+J on Psychotria M3_constrained"
plot_BioGeoBEARS_results(resDIVALIKEj, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resDIVALIKEj, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)
dev.off()


#######################################################
# BAYAREALIKE AND BAYAREALIKE+J ANALYSIS
#######################################################

# --- Run BAYAREALIKE ---
bgb_bay <- setup_base_bgb_M3(trfn, geogfn, max_range_size)

bgb_bay$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_bay$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_bay$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_bay$BioGeoBEARS_model_object@params_table["v", "type"] <- "fixed"
bgb_bay$BioGeoBEARS_model_object@params_table["v", "init"] <- 0.0
bgb_bay$BioGeoBEARS_model_object@params_table["v", "est"] <- 0.0

bgb_bay$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "1-j"
bgb_bay$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/1"
bgb_bay$BioGeoBEARS_model_object@params_table["y", "type"] <- "1-j"

bgb_bay$BioGeoBEARS_model_object@params_table["mx01y", "type"] <- "fixed"
bgb_bay$BioGeoBEARS_model_object@params_table["mx01y", "init"] <- 0.9999
bgb_bay$BioGeoBEARS_model_object@params_table["mx01y", "est"] <- 0.9999

check_BioGeoBEARS_run(bgb_bay)

resfn_bay <- "Psychotria_BAYAREALIKE_M3_constrained_v1.Rdata"
if (!file.exists(resfn_bay) | force_recalculate) {
  resBAYAREALIKE <- bears_optim_run(bgb_bay)
  save(resBAYAREALIKE, file=resfn_bay)
} else {
  load(resfn_bay)
}

# --- Run BAYAREALIKE+J ---
bgb_bayj <- setup_base_bgb_M3(trfn, geogfn, max_range_size)

dstart <- resBAYAREALIKE$outputs@params_table["d","est"]
estart <- resBAYAREALIKE$outputs@params_table["e","est"]

bgb_bayj$BioGeoBEARS_model_object@params_table["d", "init"] <- dstart
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "est"] <- dstart
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "init"] <- estart
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "est"] <- estart

bgb_bayj$BioGeoBEARS_model_object@params_table["s", "type"] <- "fixed"
bgb_bayj$BioGeoBEARS_model_object@params_table["s", "init"] <- 0.0
bgb_bayj$BioGeoBEARS_model_object@params_table["s", "est"] <- 0.0

bgb_bayj$BioGeoBEARS_model_object@params_table["v", "type"] <- "fixed"
bgb_bayj$BioGeoBEARS_model_object@params_table["v", "init"] <- 0.0
bgb_bayj$BioGeoBEARS_model_object@params_table["v", "est"] <- 0.0

bgb_bayj$BioGeoBEARS_model_object@params_table["ysv", "type"] <- "1-j"
bgb_bayj$BioGeoBEARS_model_object@params_table["ys", "type"] <- "ysv*1/1"
bgb_bayj$BioGeoBEARS_model_object@params_table["y", "type"] <- "1-j"

bgb_bayj$BioGeoBEARS_model_object@params_table["mx01y", "type"] <- "fixed"
bgb_bayj$BioGeoBEARS_model_object@params_table["mx01y", "init"] <- 0.9999
bgb_bayj$BioGeoBEARS_model_object@params_table["mx01y", "est"] <- 0.9999

bgb_bayj$BioGeoBEARS_model_object@params_table["j", "type"] <- "free"
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "init"] <- 0.0001
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "est"] <- 0.0001
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "max"] <- 0.99999

bgb_bayj$BioGeoBEARS_model_object@params_table["d", "min"] <- 0.0000001
bgb_bayj$BioGeoBEARS_model_object@params_table["d", "max"] <- 4.9999999
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "min"] <- 0.0000001
bgb_bayj$BioGeoBEARS_model_object@params_table["e", "max"] <- 4.9999999
bgb_bayj$BioGeoBEARS_model_object@params_table["j", "min"] <- 0.00001

check_BioGeoBEARS_run(bgb_bayj)

resfn_bayj <- "Psychotria_BAYAREALIKE+J_M3_constrained_v1.Rdata"
if (!file.exists(resfn_bayj) | force_recalculate) {
  resBAYAREALIKEj <- bears_optim_run(bgb_bayj)
  save(resBAYAREALIKEj, file=resfn_bayj)
} else {
  load(resfn_bayj)
}

pdf("Psychotria_BAYAREALIKE_vs_BAYAREALIKE+J_M3_constrained_v1.pdf", width=8.5, height=11)
analysis_titletxt ="BioGeoBEARS BAYAREALIKE on Psychotria M3_constrained"
plot_BioGeoBEARS_results(resBAYAREALIKE, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resBAYAREALIKE, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

analysis_titletxt ="BioGeoBEARS BAYAREALIKE+J on Psychotria M3_constrained"
plot_BioGeoBEARS_results(resBAYAREALIKEj, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)
plot_BioGeoBEARS_results(resBAYAREALIKEj, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)
dev.off()

#########################################################################
# CALCULATE SUMMARY STATISTICS TO COMPARE
#########################################################################

restable <- NULL
teststable <- NULL

# Extract parameters function
extract_params <- function(res_obj) {
  extract_params_from_BioGeoBEARS_results_object(results_object=res_obj, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
}

# --- DEC vs DEC+J ---
LnL_dec <- get_LnL_from_BioGeoBEARS_results_object(resDEC)
LnL_decj <- get_LnL_from_BioGeoBEARS_results_object(resDECj)
stats_dec <- AICstats_2models(LnL_decj, LnL_dec, 3, 2)
restable <- rbind(restable, extract_params(resDEC), extract_params(resDECj))
teststable <- rbind(teststable, conditional_format_table(stats_dec))

# --- DIVALIKE vs DIVALIKE+J ---
LnL_diva <- get_LnL_from_BioGeoBEARS_results_object(resDIVALIKE)
LnL_divaj <- get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEj)
stats_diva <- AICstats_2models(LnL_divaj, LnL_diva, 3, 2)
restable <- rbind(restable, extract_params(resDIVALIKE), extract_params(resDIVALIKEj))
teststable <- rbind(teststable, conditional_format_table(stats_diva))

# --- BAYAREALIKE vs BAYAREALIKE+J ---
LnL_bay <- get_LnL_from_BioGeoBEARS_results_object(resBAYAREALIKE)
LnL_bayj <- get_LnL_from_BioGeoBEARS_results_object(resBAYAREALIKEj)
stats_bay <- AICstats_2models(LnL_bayj, LnL_bay, 3, 2)
restable <- rbind(restable, extract_params(resBAYAREALIKE), extract_params(resBAYAREALIKEj))
teststable <- rbind(teststable, conditional_format_table(stats_bay))

# Format tables
teststable$alt <- c("DEC+J", "DIVALIKE+J", "BAYAREALIKE+J")
teststable$null <- c("DEC", "DIVALIKE", "BAYAREALIKE")
row.names(restable) <- c("DEC", "DEC+J", "DIVALIKE", "DIVALIKE+J", "BAYAREALIKE", "BAYAREALIKE+J")

# Print results
print(restable)
print(teststable)

# Save tables
save(restable, file="restable_M3_constrained_v1.Rdata")
save(teststable, file="teststable_M3_constrained_v1.Rdata")

```


Comparamos todos los modelos.

```{r}
################################################################################
# GLOBAL MODEL COMPARISON (AIC WEIGHTS)
# Comparing M0, M1, M2, and M3 settings
################################################################################

# Load the required function from BioGeoBEARS (if not already loaded)
library(BioGeoBEARS)

# 1. Load the results tables from the previous runs

# M0 (Unconstrained)
load("restable_M0_unconstrained_v1.Rdata")
restable_M0 <- restable
rownames(restable_M0) <- paste0("M0_", rownames(restable_M0)) # Add prefix

# M1 (Constrained max range = 2)
load("restable_M1_constrained_v1.Rdata")
restable_M1 <- restable
rownames(restable_M1) <- paste0("M1_", rownames(restable_M1))

# M2 (Stepping Stones)
load("restable_M2_constrained_v1.Rdata")
restable_M2 <- restable
rownames(restable_M2) <- paste0("M2_", rownames(restable_M2))

# M3 (Stratified)
load("restable_M3_constrained_v1.Rdata") # Make sure the M3 file is correctly named here
restable_M3 <- restable
rownames(restable_M3) <- paste0("M3_", rownames(restable_M3))

# 2. Combine all tables into a single master table (24 rows)
master_restable <- rbind(restable_M0, restable_M1, restable_M2, restable_M3)

# 3. Calculate AIC for all models
# AIC = 2*k - 2*LnL  (where k is the number of parameters)
AIC_values <- 2 * master_restable$numparams - 2 * master_restable$LnL
master_restable$AIC <- AIC_values

# 4. Calculate AIC differences (Delta AIC)
# The best model has the lowest AIC. Delta AIC is the difference between each model and the best one.
min_AIC <- min(AIC_values)
master_restable$deltaAIC <- AIC_values - min_AIC

# 5. Calculate Relative Likelihoods of the models
# exp(-0.5 * deltaAIC)
rel_likes <- exp(-0.5 * master_restable$deltaAIC)

# 6. Calculate AIC Weights
# AIC weight represents the probability that a given model is the best among the set of candidate models.
master_restable$AIC_weight <- rel_likes / sum(rel_likes)

# 7. Sort the table from best model (highest weight) to worst
master_restable_sorted <- master_restable[order(master_restable$AIC_weight, decreasing = TRUE), ]

# Print the final formatted table
print(conditional_format_table(master_restable_sorted))

# Save the final comparison
save(master_restable_sorted, file="Global_Model_Comparison_M0_M1_M2_M3.Rdata")
write.table(master_restable_sorted, file="Global_Model_Comparison_M0_M1_M2_M3.txt", sep="\t", quote=FALSE)
```



Finalmente, vamos a hacer mapeo estocástico usando el mejor modelo inferido (según AIC es el modelo DEC+J y con steeping stones)

```{r}

################################################################################
# STOCHASTIC MAPPING (BSM) ON THE BEST INFERRED MODEL
# Winning Model: DEC+J under M2 (Stepping Stones)
################################################################################

library(BioGeoBEARS)

# 1. Load the winning model directly (No need to recalculate!)
resfn_decj <- "Psychotria_DEC+J_M2_constrained_v1.Rdata"
load(resfn_decj)

# BioGeoBEARS stochastic mapping functions expect the model object to be named 'res'
res <- resDECj 

#######################################################
# Get the inputs for Biogeographical Stochastic Mapping
#######################################################
BSM_inputs_fn = "BSM_inputs_file.Rdata"
runInputsSlow = TRUE

if (runInputsSlow) {
    stochastic_mapping_inputs_list = get_inputs_for_stochastic_mapping(res=res)
    save(stochastic_mapping_inputs_list, file=BSM_inputs_fn)
} else {
    load(BSM_inputs_fn)
}

# Set seed for reproducibility
set.seed(seed=as.numeric(Sys.time()))

#######################################################
# Run the Biogeographical Stochastic Mapping (50 maps)
#######################################################
runBSMslow = TRUE

if (runBSMslow == TRUE) {
    # Generate 50 stochastic maps
    BSM_output = runBSM(res, stochastic_mapping_inputs_list=stochastic_mapping_inputs_list, maxnum_maps_to_try=100, nummaps_goal=50, maxtries_per_branch=40000, save_after_every_try=TRUE, savedir=getwd(), seedval=12345, wait_before_save=0.01, master_nodenum_toPrint=0)
    
    RES_clado_events_tables = BSM_output$RES_clado_events_tables
    RES_ana_events_tables = BSM_output$RES_ana_events_tables
} else {
    # Load previously saved maps
    load(file="RES_clado_events_tables.Rdata")
    load(file="RES_ana_events_tables.Rdata")
    BSM_output = list()
    BSM_output$RES_clado_events_tables = RES_clado_events_tables
    BSM_output$RES_ana_events_tables = RES_ana_events_tables
}

# Extract BSM output tables
clado_events_tables = BSM_output$RES_clado_events_tables
ana_events_tables = BSM_output$RES_ana_events_tables

#######################################################
# Setup colors and states for plotting
#######################################################
include_null_range = TRUE
areanames = names(tipranges@df)
areas = areanames
max_range_size = 2

states_list_0based = rcpp_areas_list_to_states_list(areas=areas, maxareas=max_range_size, include_null_range=include_null_range)
colors_list_for_states = get_colors_for_states_list_0based(areanames=areanames, states_list_0based=states_list_0based, max_range_size=max_range_size, plot_null_range=TRUE)

scriptdir = normalizePath(system.file("extdata/a_scripts", package="BioGeoBEARS"))
stratified = FALSE # M2 is spatial constraint, not time-stratified

#######################################################
# Plot a single stochastic map
#######################################################
pdffn_single = "Psychotria_DEC+J_M2_constrained_single_stochastic_map_n1.pdf"
pdf(file=pdffn_single, height=6, width=6)

master_table_cladogenetic_events = clado_events_tables[[1]]
resmod = stochastic_map_states_into_res(res=res, master_table_cladogenetic_events=master_table_cladogenetic_events, stratified=stratified)

plot_BioGeoBEARS_results(results_object=resmod, analysis_titletxt="Stochastic map", addl_params=list("j"), label.offset=0.5, plotwhat="text", cornercoords_loc=scriptdir, root.edge=TRUE, colors_list_for_states=colors_list_for_states, skiptree=FALSE, show.tip.label=TRUE)

paint_stochastic_map_branches(res=resmod, master_table_cladogenetic_events=master_table_cladogenetic_events, colors_list_for_states=colors_list_for_states, lwd=5, lty=par("lty"), root.edge=TRUE, stratified=stratified)

plot_BioGeoBEARS_results(results_object=resmod, analysis_titletxt="Stochastic map", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, root.edge=TRUE, colors_list_for_states=colors_list_for_states, skiptree=TRUE, show.tip.label=TRUE)

dev.off()

#######################################################
# Plot all 50 stochastic maps
#######################################################
pdffn_all = paste0("Psychotria_DEC+J_M2_constrained_", length(clado_events_tables), "BSMs.pdf")
pdf(file=pdffn_all, height=6, width=6)

nummaps_goal = 50
for (i in 1:nummaps_goal) {
    clado_events_table = clado_events_tables[[i]]
    analysis_titletxt = paste0("DEC+J M2 - Stochastic Map #", i, "/", nummaps_goal)
    plot_BSM(results_object=res, clado_events_table=clado_events_table, stratified=stratified, analysis_titletxt=analysis_titletxt, addl_params=list("j"), label.offset=0.5, plotwhat="text", cornercoords_loc=scriptdir, root.edge=TRUE, colors_list_for_states=colors_list_for_states, show.tip.label=TRUE, include_null_range=include_null_range)
}
dev.off()

#######################################################
# Summarize stochastic map tables & counts
#######################################################
actual_names = areanames

# Simulate the source areas
BSMs_w_sourceAreas = simulate_source_areas_ana_clado(res, clado_events_tables, ana_events_tables, areanames)
clado_events_tables = BSMs_w_sourceAreas$clado_events_tables
ana_events_tables = BSMs_w_sourceAreas$ana_events_tables

# Count all anagenetic and cladogenetic events
counts_list = count_ana_clado_events(clado_events_tables, ana_events_tables, areanames, actual_names)

summary_counts_BSMs = counts_list$summary_counts_BSMs
print(conditional_format_table(summary_counts_BSMs))

# Histogram of event counts
hist_event_counts(counts_list, pdffn="Psychotria_DEC+J_M2_constrained_histograms_of_event_counts.pdf")

#######################################################
# Print counts to files
#######################################################
tmpnames = names(counts_list)
cat("\n\nWriting tables of counts to tab-delimited text files...\n\n")
for (i in 1:length(tmpnames)) {
    cmdtxt = paste0("item = counts_list$", tmpnames[i])
    eval(parse(text=cmdtxt))

    # Skip cubes (3D matrices)
    if (length(dim(item)) != 2) next()

    outfn = paste0("Psychotria_DEC+J_M2_constrained_", tmpnames[i], ".txt")
    if (length(item) == 0) {
        cat(outfn, " -- NOT written, *NO* events recorded of this type\n", sep="")
    } else {
        write.table(conditional_format_table(item), file=outfn, quote=FALSE, sep="\t", col.names=TRUE, row.names=TRUE)
        cat(outfn, " -- Written.\n", sep="")
    }
}
cat("...done.\n")

#######################################################
# Check ML vs BSM consistency
#######################################################
library(MultinomialCI)    # Required for 95% CIs on BSM counts
check_ML_vs_BSM(res, clado_events_tables, "DEC+J_M2_constrained", tr=NULL, plot_each_node=FALSE, linreg_plot=TRUE, MultinomialCI=TRUE)

```




La bibliografía recomendada es la misma que para la sesión 11, pero vamos a añadir algunas referencias adicionales.

BIBLIOGRAFÍA ADICIONAL

Hackel, J., Vorontsova, M. S., Nanjarisoa, O. P., Hall, R. C., Razanatsoa, J., Malakasi, P., & Besnard, G. (2018). Grass diversification in Madagascar: in situ radiation of two large C3 shade clades and support for a Miocene to Pliocene origin of C4 grassy biomes. Journal of Biogeography, 45(4), 750-761.

Larridon, I., Galan Diaz, J., Bauters, K., & Escudero, M. (2021). What drives diversification in a pantropical plant lineage with extraordinary capacity for long‐distance dispersal and colonization?. Journal of Biogeography, 48(1), 64-77.

Larridon, I., Spalink, D., Jiménez‐Mejías, P., Márquez‐Corro, J. I., Martín‐Bravo, S., Muasya, M., & Escudero, M. (2021). The evolutionary history of sedges (Cyperaceae) in Madagascar. Journal of Biogeography, 48(4), 917-932.

Maguilla, E., Escudero, M., & Luceño, M. (2018). Vicariance versus dispersal across Beringian land bridges to explain circumpolar distribution: A case study in plants with high dispersal potential. Journal of Biogeography, 45(4), 771-783.

Maguilla, E., Escudero, M., Jiménez-Lobato, V., Díaz-Lifante, Z., Andrés-Camacho, C., & Arroyo, J. (2021). Polyploidy expands the range of Centaurium (Gentianaceae). Frontiers in Plant Science, 12, 650551.

Maguilla, E., Escudero, M., Ruíz‐Martín, J., & Arroyo, J. (2021). Origin and diversification of flax and their relationship with heterostyly across the range. Journal of Biogeography, 48(8), 1994-2007.


Además también la bibliografía de la sesión anterior:

BIBLIOGRAFÍA BÁSICA

Ree, R. H., and S. A. Smith. 2008. Maximum likelihood inference of geographic range evolution by
dispersal, local extinction, and cladogenesis. Syst. Biol. 57:4–14.

BIBLIOGRAFÍA RECOMENDADA

Dupin, J., Matzke, N. J., Särkinen, T., Knapp, S., Olmstead, R. G., Bohs, L., & Smith, S. D. (2017). Bayesian estimation of the global biogeographical history of the Solanaceae. Journal of Biogeography, 44(4), 887-899.

Klaus, K. V., and N. J. Matzke. 2020. Statistical comparison of trait-dependent biogeographical models
indicates that podocarpaceae dispersal is influenced by both seed cone traits and geographical
distance. Syst. Biol. 69:61–75.

Matzke, N. J. 2012. Founder-event speciation in BioGeoBEARS package dramatically improves likeli-
hoods and alters parameter inference in dispersal-extinction-cladogenesis (DEC) analyses. Front.
Biogeogr 4:210.

Matzke, N. J. 2013. Probabilistic historical biogeography: New models for founder-event speciation, imperfect detection, and fossils allow improved accuracy and model-testing. Front. Biogeogr. 5:242–48.

Ree, R. H., B. R. Moore, C. O. Webb, and M. J. Donoghue. 2005. A likelihood framework for inferring the evolution of geographic range on phylogenetic trees. Evolution 59:2299–311.

Ree, R. H., and I. Sanmartin. 2018. Conceptual and statistical problems with the DEC+J model of
founder-event speciation and its comparison with DEC via model selection. J. Biogeogr. 45:
741–49.

Van Dam, M. H., & Matzke, N. J. (2016). Evaluating the influence of connectivity and distance on biogeographical patterns in the south‐western deserts of North America. Journal of Biogeography, 43(8), 1514-1532.